\chapter{dataReset}

\section{Introduction}

A linearized process may have parameters that are initialized, changed and used, and subsequently ignored until they are used again.
In the interval from the last change or use of a parameter to its subsequent reuse, a parameter could have different values.
These values contribute to the size of the state space of the process \emph{without} adding any new behavior!

It may be advantageous to detect from which moment the value of a parameter is no longer used and set it to a default value instead.
This is called `resetting'.
Analysis of an LPE can yield information on when parameters can be safely reset.

\section{Algorithm}

The algorithm is based on earlier work \cite{van2009state}.

As a preparative step, several determinations are made about each parameter $p$ of the LPE per summand $s$:
\begin{itemize}

\item Is the value of $p$ potentially altered by $s$?
If so, $p$ is marked as `changed'.

\item Does $p$ occur in the guard of $s$?
If so, $p$ is marked as `directly used'.

\item Is $p$ marked as `directly used' in $s$?
Or does $p$ occur in the assignment by $s$ to a parameter that is marked as `changed'?
In either case, $p$ is marked as `used'.

\item Must the value of $p$ have a specific (unique) value $v$ in order for $s$ to be enabled?
If so, $p$ is marked as `having a source' for $s$, with the source being $v$.

\item Does the value of $p$ have a specific (unique) value $v$ immediately after $s$ has been applied?
If so, $p$ is marked as `having a destination' for $s$, with the source being $v$.

\item Does $p$ have both a source and a destination for $s$?
Then $p$ is called a \emph{ruling parameter}.

\end{itemize}

\subsection{Control flow graphs}

A parameter of the LPE is a \emph{control flow parameter} of the LPE if for all summands it is either a ruling parameter or not marked as `changed' (or both).
A parameter that is not a control flow parameter is defined as a \emph{data parameter}.

\vspace{1mm}

For each control flow parameter $f$ of the LPE, a control flow graph is constructed.
There is a state in the control flow graph for each source and destination of $f$ (across all summands).
Two states $s_1$ and $s_2$ are connected by an edge $(s_1, i, s_2)$ if there is a summand $i$ where the source of $f$ is represented by one of the states and where the destination of $f$ is represented by the other state.
The direction of such an edge is from source state to destination state.

\subsection{Belongs-to function}

Next, we define the \emph{belongs-to} function.
The belongs-to function maps each data parameter $d$ to some set of control flow parameters $b(d)$ as follows

\begin{align*}
b(d) = F \cap \bigcap\limits_{s \in S}^{} \text{ruling}(s)
\end{align*}

where

\begin{itemize}
\item $F$ is the set of all control flow parameters;
\item $S$ is the set of all summands in which $d$ is marked as `changed' or `used';
\item $\text{ruling}(s)$ is the set of all parameters that rule summand $s$.
\end{itemize}

\clearpage
\subsection{Relevance relation}

The relevance relation $R$ bundles the information that has been gathered so far.
The relation $R$ relates a data parameter $d$, a control flow parameter $f$, and a value $v$ if $d$ may be used after a state in which $f$ has value $v$.
This is denoted $R(d, f, v)$.
Intuitively, this is a situation in which $d$ should \emph{not} be reset.

The computation of $R$ is a fixpoint algorithm: modifications are applied iteratively until $R$ no longer changes.
The initial value of $R$ is set to

\begin{align*}
R_0 = \bigcup\limits_{\substack{i \in S \\ d_k \in \text{directlyUsed}(i)}}^{} \;\{\; (d_k, d_j, \text{source}(i, d_j)) \;|\; d_j \in b(d_k) \;\}
\end{align*}

where

\begin{itemize}
\item $S$ is the set of all summands;
\item $\text{directlyUsed}(i)$ is a function that gives the set of all parameters marked as `directly used' in a summand $i$;
\item $\text{source}(i, f)$ is a function that gives the source value of a control flow parameter $f$ for a summand $i$.
\end{itemize}

Each iteration can be split into two steps.
The first step checks for control flow graphs in which a data parameter $d$ has already been marked as `relevant' whether this implies that $d$ is also relevant in preceding states (of the same control flow graph).
If so, the appropriate triples are added to $R$:

\begin{align*}
R_{n}{'} = R_{n-1} \cup \bigcup\limits_{\substack{i \in S}}^{} \;\left\{\; (d_k, d_j, s) \;\middle|\; \substack{(d_l, d_j, t) \in R_{n-1} \\ d_j \in b(d_k) \\ d_k \in \text{vars}(v_i(d_l)) \\ (s, i, t) \in E_{d_j}} \;\right\}
\end{align*}

where

\begin{itemize}
\item $E_{f}$ is the set of edges that are part of the control flow graph of control flow parameter $f$.
\end{itemize}

The second step is similar, but data parameters that are found to be relevant are added as triples to $R$ in relation to \emph{another} control flow parameter:

\begin{align*}
R_{n} = R_{n}{'} \cup \bigcup\limits_{\substack{i \in S}}^{} \;\left\{\; (d_k, d_j, \text{source}(i, d_j)) \;\middle|\; \substack{(d_l, d_p, t) \in R_{n}{'} \\ d_j \in b(d_k),\; d_j \notin b(d_l) \\ d_k \in \text{vars}(v_i(d_l)) \\ (r, i, t) \in E_{d_p}} \;\right\}
\end{align*}

\clearpage
\subsection{Summand modification}

Finally, each summand of an LPE can be modified by making use of the belongs-to function $b$ and the relevance relation $R$.
Intuitively, we check whether the value of a parameter is `relevant' after a specific summand $i$ has been applied.
If so, the parameter should not be reset.
Otherwise, the parameter can be reset, for example to its initialization value:

\begin{align*}
v_{i}{'}(d_k) = \begin{cases}
v_{i}(d_k) & \text{if } \bigwedge\limits_{\substack{d_j \in \text{ruling}(i) \\ d_j \in b(d_k)}}^{} R(d_k, d_j, \text{dest}(i, d_j)) \\
v_{0}(d_k) & \text{otherwise}
\end{cases}
\end{align*}

where

\begin{itemize}
\item $\text{ruling}(s)$ is the set of all parameters that rule summand $s$;
\item $\text{source}(i, f)$ is a function that gives the destination value of a control flow parameter $f$ for a summand $i$.
\end{itemize}

